---
title: "Temp_Time_Series"
author: "Sam Bogan"
date: "6/23/2021"
output: github_document
---

This is an R Markdown written by Sam Bogan that reads in, wrangles, summarizes, and analyzes time series temperatures collected from n = 3 per site supralittoral splash pools occupied by *Tigriopus californicus* across 4 sites in coastal California (see README). Data were recorded by TidBit MX5000 temperature loggers made by Onset Computer Corp between 2019 - 2021. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Identify measurements to filter out

```{r}

# Load packages
library( tidyverse )
library( Rmisc )
library( lubridate )
library( data.table )
library( tibbletime )

# Read csv describing periods where loggers broke or were aerially exposed
TidBit_meta_df <- read.csv( "~/Documents/GitHub/HotOnes_Tigriopus/Temp_Time_Series/Input_files/TidBit_Deployment_Check_History.csv" )

# Print metadata sheet
TidBit_meta_df

## Filter out periods where logger may have been aerially exposed due to evaporation of pool
# Start by outputing rows in metadata signifying post-date filtering
filt_post_dates <- filter( TidBit_meta_df, Filter_Post_Date == "Yes" )

# Print dates
filt_post_dates

# Create variables for filtering
logger_to_filter <- as.character( filt_post_dates$Logger )
date_to_filter_past <- as.Date( filt_post_dates$Date, "%m/%d/%y" )

```

#Read and wrangle raw data

```{r}

# Move to input data directory for tempm time series
setwd( "~/Documents/GitHub/HotOnes_Tigriopus/Temp_Time_Series/Input_files/" )

# Get the files names
temps <- list.files( pattern = "*0.csv" )

# Read in .csv files listed in temps
mytemps <- lapply( temps, read.csv )

# Correct df names from mytemps
names( mytemps ) <- gsub( " .*",
                          "",
                          temps )

# Remove unecessary columns and standardize column number
mytemps <- lapply( mytemps, function ( y ) { y <- select( y, c( 1, 2 ) ) } )

# Filter out contiously aerially-exposed measurements in BRM2
mytemps$BMR2 <- dplyr::filter( mytemps$BMR2, as.Date( Date.Time..GMT..0800, "%m/%d/%y" ) < as.Date( date_to_filter_past ) )

# Data from PTD3 must be filtered following 02/02/2021 due to damage and miscalibration
mytemps$PTD3 <- dplyr::filter( mytemps$PTD3, as.Date(  mytemps$PTD3$Date.Time..GMT..0700, "%m/%d/%y" ) < as.Date( "2021-02-02 00:00:00" ) )

# Combine all temp dfs
all_temps <- rbindlist( mytemps, 
                        idcol = "Logger" )

# Remove rows where temp is empty
all_temps <- all_temps[ !(is.na( all_temps$Temp....F. ) | 
                            all_temps$Temp....F.  == "" ), ]

# Create site variable
all_temps$Site <- gsub( "[1-9]+", 
                        "", 
                        all_temps$Logger )

# Fix column names
names( all_temps )[ names( all_temps ) == "Date.Time..GMT..0800"] <- "Date_Time"
names( all_temps )[ names( all_temps ) == "Temp....F."] <- "Temp"

# Convert temp to celcius
all_temps$Temp <- round( ( ( all_temps$Temp - 32 ) * ( 5 / 9 ) ),
                         digits = 2 )

# Create date variable
all_temps$Date <- as.Date( all_temps$Date_Time, "%m/%d/%y" )

# Ensure that Date_Time variable is in date-time format
all_temps$Date_Time <- mdy_hm( all_temps$Date_Time )

# Create month variable
all_temps$Month <- month( all_temps$Date )

# Create latitude index
lat_index <- data.frame( Site = c( "BMR", "SC", "RMR", "PTD" ),
                         Lat = c( 38.316375,
                                  36.965681,
                                  35.540147,
                                  34.002089 ) )
# Merge lat index with temp df
all_temps <- merge( all_temps,
       lat_index,
       by = "Site" )

```

#Plot filtered temperatures

```{r}

# Plot all temp data as check on filtering and wrangling
ggplot( data = all_temps, 
        aes( x = Date_Time, y = Temp ) ) +
  geom_line( alpha = 0.5, size = 0.5, stat = "identity" ) +
  theme_classic() +
  facet_wrap( ~Logger )

# Plot all temps by site
ggplot( data = all_temps, 
        aes( x = Date_Time, y = Temp, group = Site ) ) +
  geom_point( alpha = 0.025, size = 0.05 ) +
  theme_gray() +
  facet_grid( Site ~ . )

# Filter to common earliest and latest dates for all loggers before plott means: one day after PTD deployment
all_temps_filt <- filter( all_temps, 
                          Date > as.Date( "2020-03-01" ),
                          Date < as.Date( "2021-06-12" ) )

all_temps_filt_1yr <- filter( all_temps, 
                          Date > as.Date( "2020-06-12" ),
                          Date < as.Date( "2021-06-12" ) )

# Estimate daily means
mean_temps <- summarySE( measurevar = "Temp",
                         groupvars = c( "Site", "Date", "Lat", "Month" ),
                         data = all_temps_filt )

# Convert mean date to datetime variable
names( mean_temps )[ names( mean_temps ) == "Date"] <- "Date_Time"
mean_temps$Date_Time <- as_datetime( mean_temps$Date_Time )

# Estimate overall annual mean
overall_mean_temp <- summarySE( measurevar = "Temp",
                         groupvars = c( "Site", "Date" ),
                         data = all_temps_filt_1yr )

overall_mean_temp <- summarySE( measurevar = "Temp",
                         groupvars = NULL,
                         data = overall_mean_temp )

#Print overall mean temp
overall_mean_temp

# Estimate site means
site_temps <- summarySE( measurevar = "Temp",
                         groupvars = c( "Site", "Lat" ),
                         data = all_temps_filt )

# Estimate site means by month
site_month_temps <- summarySE( measurevar = "Temp",
                         groupvars = c( "Site", "Lat", "Month" ),
                         data = all_temps_filt )

# Plot daily means over time, colored by site
ggplot( data = mean_temps, 
        aes( x = Date_Time, y = Temp, 
             group = Site, color = -Lat, fill = -Lat ) ) +
  geom_smooth( method = "loess", se = T ) +
  scale_color_viridis_c() +
  scale_fill_viridis_c( guide = FALSE ) +
  labs( x = "Date", y = "Temp (°C)" , color = "Latitude (°N)" ) +
  theme_classic()

# Plot distributions of daily mean temp, grouped by site
ggplot( data = mean_temps, 
        aes( x = Lat, y = Temp, group = Site ) ) +
  geom_jitter( size = 0.5, width = 0.1, color = "grey" ) +
  geom_point( data = site_temps, size = 3 ) +
  geom_errorbar( data = site_temps,
                 aes( ymin = Temp - ci, ymax = Temp + ci ),
                 width = 0, size = 1 ) +
  labs( y = "Temp (°C)" , x = "Latitude (°N)" ) +
  geom_line( data = site_temps, group = "identity" ) +
  theme_classic()

# Plot relationship between latitude and temp by month and year: latitudinal temp gradient only holds in winter months
ggplot( data = mean_temps, 
        aes( x = Lat, y = Temp, group = Site ) ) +
  geom_jitter( size = 0.5, width = 0.1, color = "grey" ) +
  geom_line( data = site_month_temps, group = "identity" ) +
  geom_point( data = site_month_temps, size = 3, aes( color = Temp ) ) +
  facet_wrap( ~Month, scale = "free" ) +
  scale_color_viridis_c() +
  labs( y = "Temp (°C)", x = "Latitude (°N)", color = "Temp (°C)" ) +
  theme_classic() +
  theme( strip.background = element_blank() )


```


